---
title: "Exercise 3"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#install.packages('arrow')
#install.packages('gender')
#install.packages('genderdata')
library("arrow")
library("tidyverse")
library("gender")

```

## Load the files  

You can include R code in the document as follows:

```{r}
# change to your own path!
data_path <- "/Users/tehuttesfayebiru/projects/2022-ona-assignments/exercises/"
applications <- read_parquet(paste0(data_path,"app_data_sample.parquet"))
edges <- read_csv(paste0(data_path,"edges_sample.csv"))

```
```{r}
applications
```
```{r}
edges
```
## add the following variables for examiners: Gender, Race, Tenure


## Gender
```{r}
# get a list of first names without repetitions
# get a list of first names without repetitions
examiner_names <- applications %>% 
  distinct(examiner_name_first)

examiner_names
```

```{r}
# get a table of names and gender
examiner_names_gender <- examiner_names %>% 
  do(results = gender(.$examiner_name_first, method = "ssa")) %>% 
  unnest(cols = c(results), keep_empty = TRUE) %>% 
  select(
    examiner_name_first = name,
    gender,
    proportion_female
  )

examiner_names_gender
```

```{r}
# remove extra colums from the gender table
examiner_names_gender <- examiner_names_gender %>% 
  select(examiner_name_first, gender)

# joining gender back to the dataset
applications <- applications %>% 
  left_join(examiner_names_gender, by = "examiner_name_first")

# cleaning up
rm(examiner_names)
rm(examiner_names_gender)
gc()

applications
```

## Race
```{r}
library(wru)

examiner_surnames <- applications %>% 
  select(surname = examiner_name_last) %>% 
  distinct()

examiner_surnames
```

```{r}
examiner_race <- predict_race(voter.file = examiner_surnames, surname.only = T) %>% 
  as_tibble()
examiner_race
```

```{r}
examiner_race <- examiner_race %>% 
  mutate(max_race_p = pmax(pred.asi, pred.bla, pred.his, pred.oth, pred.whi)) %>% 
  mutate(race = case_when(
    max_race_p == pred.asi ~ "Asian",
    max_race_p == pred.bla ~ "black",
    max_race_p == pred.his ~ "Hispanic",
    max_race_p == pred.oth ~ "other",
    max_race_p == pred.whi ~ "white",
    TRUE ~ NA_character_
  ))

examiner_race
```

```{r}
# removing extra columns
examiner_race <- examiner_race %>% 
  select(surname,race)

applications <- applications %>% 
  left_join(examiner_race, by = c("examiner_name_last" = "surname"))

rm(examiner_race)
rm(examiner_surnames)
gc()
```


```{r}
applications
```

## Tenure

```{r}
library(lubridate) # to work with dates

examiner_dates <- applications %>% 
  select(examiner_id, filing_date, appl_status_date) 

examiner_dates
```

```{r}
examiner_dates <- examiner_dates %>% 
  mutate(start_date = ymd(filing_date), end_date = as_date(dmy_hms(appl_status_date)))
```

```{r}
examiner_dates <- examiner_dates %>% 
  group_by(examiner_id) %>% 
  summarise(
    earliest_date = min(start_date, na.rm = TRUE), 
    latest_date = max(end_date, na.rm = TRUE),
    tenure_days = interval(earliest_date, latest_date) %/% days(1)
    ) %>% 
  filter(year(latest_date)<2018)

examiner_dates
```


```{r}
applications <- applications %>% 
  left_join(examiner_dates, by = "examiner_id")

rm(examiner_dates)
gc()

applications
```

## Pick two workgroups you want to focus on (remember that a workgroup is
## represented by the first 3 digits of `examiner_art_unit` value)

How do they compare on examiners’ demographics? Show summary statistics and
plots.

```{r}

group_176 = applications[substr(applications$examiner_art_unit, 1,3)==176,]
group_162 = applications[substr(applications$examiner_art_unit, 1,3)==162,]
summary(group_176)
```

```{r}
summary(group_162)
```
- Tenure for work group 162 seems to be a bit higher
- There are many more examiers in work group 162
# histograms for gender and race

```{r}
library(gridExtra)
par(mfrow=c(1,2))
plot1 <- ggplot(group_162, aes(x = factor(gender))) +
    geom_bar(fill="darkblue")  
plot2 <- ggplot(group_176, aes(x = factor(gender))) +
    geom_bar(fill="darkred")
grid.arrange(plot1, plot2, ncol=2)
```
For work group 162, the ratio between male and female examiners seems to be somewhat even, but there are a lot of examiners where the gender is unknown compared to group 176. 

For work group 176, there are many more male examiners compared to female examiners. 
```{r}
library(gridExtra)
par(mfrow=c(1,2))
plot1 <- ggplot(group_162, aes(x = factor(race))) +
    geom_bar(fill="darkblue")  
plot2 <- ggplot(group_176, aes(x = factor(race))) +
    geom_bar(fill="darkred")
grid.arrange(plot1, plot2, ncol=2)
```

From both work groups, we see that the number of white examiners greatly outnumbers all other races. The second most prominent race group is asian. While both these racial groups are relatively even for both working groups, working group 162 has more black examiners. 


## 3. Create advice networks from `edges_sample` and calculate centrality scores
## for examiners in your selected workgroups

### Pick measure(s) of centrality you want to use and justify your choice
### Characterize and discuss the relationship between centrality and other examiners’
### characteristics

```{r}
# get examiner ids to use as nodes
examiner_ids = distinct(subset(applications, select=c(examiner_art_unit, examiner_id)))
examiner_ids$workgroup = substr(examiner_ids$examiner_art_unit, 1,3)
examiner_ids = examiner_ids[examiner_ids$workgroup==162 | examiner_ids$workgroup==176,]

# merge with edges dataframe to get final edge table
data.f = merge(x=edges, y=examiner_ids, by.x="ego_examiner_id", by.y="examiner_id", all.x=TRUE)
data.f = data.f %>% rename(ego_art_unit=examiner_art_unit, ego_workgroup=workgroup)
data.f = drop_na(data.f)

data.f = merge(x=data.f, y=examiner_ids, by.x="alter_examiner_id", by.y="examiner_id", all.x=TRUE)
data.f = data.f %>% rename(alter_art_unit=examiner_art_unit, alter_workgroup=workgroup)
data.f = drop_na(data.f)

```

```{r}
# get unique ego and alter nodes
ego_nodes = subset(data.f, select=c(ego_examiner_id,ego_art_unit, ego_workgroup)) %>% rename(examiner_id=ego_examiner_id,art_unit=ego_art_unit,workgroup=ego_workgroup)
alter_nodes = subset(data.f, select=c(alter_examiner_id,alter_art_unit, alter_workgroup))%>% rename(examiner_id=alter_examiner_id,art_unit=alter_art_unit,workgroup=alter_workgroup)
nodes = rbind(ego_nodes, alter_nodes)
nodes = distinct(nodes)
nodes = nodes %>% group_by(examiner_id) %>% summarise(examiner_id=first(examiner_id), art_unit=first(art_unit), workgroup=first(workgroup))
library(igraph)
# creating network
advice_net = graph_from_data_frame(d=data.f, vertices=nodes, directed=TRUE)
```

### Calculate centrality scores

```{r}
Degree <- degree(advice_net, v=V(advice_net))
Betweenness <- betweenness(advice_net)
Closeness <- closeness(advice_net)
```

### Visualize centralities
```{r}
# save centrality scores
V(advice_net)$size = Degree
V(advice_net)$clos = round(Closeness,2)
V(advice_net)$bet = round(Betweenness,2)
# color by art unit
V(advice_net)$color = nodes$art_unit

# save scores in a data frame for easy visualization
centralities <- cbind(Degree, Closeness, Betweenness)
centralities = round(centralities,2)
centralities = data.frame(centralities)

# plot graph 
ggraph(advice_net, layout="kk") +
       geom_edge_link()+
       geom_node_point(aes(size=size, color=color), show.legend=T) 
```


We can see from this network visualization that most examiners seek advice from their own art units, with the most connections (or examiners with the largest advice networks and connections) are seen for work groups starting with the identifier 17. Interestingly, some examiners from this work group also affiliate with advice networks consisting of people from art units with identifiers starting with 16. Taking a second look at the graph with examiner id values, we see hat examiners with ID 72941 and 92836 are those which are part of advice networks in a separate art unit from their own.

```{r}
# show info for the first examiner being discussed
first_examiner <- applications %>% 
  filter(examiner_id==72941)
summary(first_examiner)
```



```{r}
# show info for the second examiner being discussed
second_examiner <- applications %>% 
  filter(examiner_id==92836)
summary(second_examiner)
```
```{r}
first_examiner
```

```{r}
second_examiner
```

Looking at these two examiners, we can see that they have both moved across different art units during their time at the USPTO.They also have both worked at this institution more than 15 years (high tenure). Both examiners are female, with one predicted as asian and the other as white. A further investigation into these and more characteristics may be able to inform us on why these examiner chose to affiliate with advice networks outside of their current art unit. 



```{r}
# a bit messier graph
ggraph(advice_net, layout="kk") +
       geom_edge_link()+
       geom_node_point(aes(size=size, color=color), show.legend=T) +
       geom_node_text(aes(label = name))
```
