---
title: "Exercise 2"
output: github_document

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading libraries

```{r}
library(dplyr)
library(tidygraph)
library(tidyverse)
library(ggraph)
```

## Creating nodes and edges

###### Nodes: [1,2,3,4,5,6,A,B,C,D]
###### Connections between 'nodes': 

6 -> 5, 6 -> D, 6 -> B, 1 -> 2, 2 -> A, A -> B, A -> C, 
B -> C, B -> D, B -> 3, C -> D, C -> 3, C -> 4, D -> 3, 
D -> 5, 3 -> 5, 3 -> 4

###### Load CSV file that contains the edge table

```{r}
bus_network <- read.csv('fb_bus.csv')
bus_network
```

### Create network graph 

```{r}
bus_graph = as_tbl_graph(bus_network)
```

### Plotting network graph

```{r}
ggraph(bus_graph) + 
  geom_edge_link() + 
  geom_node_point()+
  geom_node_text(aes(label = name)) +
  theme_graph()
```

### Calculating metrics

###### Degree centrality - how many edge connections for each node

```{r}
# add results to graph
bus_graph =bus_graph %>%
  activate(nodes) %>%
  mutate(centrality_degree = centrality_degree())

bus_graph

```

###### Betweenness centrality - using shortest paths

``` {r}
# add results to graph
bus_graph = bus_graph %>%
  activate(nodes) %>%
  mutate(centrality_betweenness = centrality_betweenness())
bus_graph
```

###### Closeness centrality - how close to all other nodes

```{r}
# add results to graph
bus_graph = bus_graph %>%
  activate(nodes) %>%
  mutate(centrality_closeness = centrality_closeness())

bus_graph
```

### Let's look at the metrics for only the nodes where it is possible to sit

``` {r}
bus_graph_options <- bus_graph %>% 
  filter(name %in% c('A','B','C','D'))

bus_graph_options
```

### Discuss possible consequences of your choice of a seat. When would this choice be beneficial? When would it be not so beneficial?

The best option to sit would be the seat that has relatively higher values in each of the metrics in comparison to the other seats. This option is C, which would give one of the highest degree centralities, the option to speak to multiple people on the bus. It also has one of the higher values for betweenness and closeness centrality, which would mean that seat would more likely serve as a bridge of communication or information sharing between other seats. However, if the main objective if to serve as a bridge in between others, the best option would have been seat A

### Plot the network graph with labels and centrality values

```{r}
#install.packages("graphlayouts")
library(graphlayouts)

```


```{r}
ggraph(bus_graph, "stress", bbox = 15) +
  geom_edge_link2(aes(edge_colour = "orange"), edge_width = 0.5) +
  geom_node_point(aes(fill = centrality_betweenness, size = centrality_degree), shape = 21) +
  geom_node_text(aes(label = name, size=centrality_closeness),
    family = "serif", repel = TRUE
  ) +
  scale_edge_colour_brewer(palette = "Set1") +
  scale_size(range = c(2, 5), guide = "none") +
  theme_graph() 
```
The above plot colors the nodes based on centrality measures, with the lightest blue nodes having the highest betweenness centrality values. Larger nodes and label sizes also show higher degree centrality and closeness centrality values respectively.

